---
title: "BNI assessment for plant root exudates demonstration"
author: "Alex Kanellopoulos"
date: "2024-08-22"
output: html_document
---

Brief description: This is a script that includes analyses for the assessment of BNI activity of different plant root exudates. In the dataset, I have included 44 and 33 root exudates from equal different genotypes of two different plant species. These exudates were screened for BNI activity against 4 AOM strains, belonging in 2 different AOM Domains, and the AOI (Ammonia Oxidation Inhibition %) is given. The analyses include descriptive statistics for the BNI activity, followed by a section for producing barplots for the different exudates across all 4 AOM strains. Finally, I have included data visualisation in a heatmap format and investigation of inhibitory profiles using a PCA. 

Step 1: Let's see all the necessary packages needed for the analysis and load them. If they are not installed, we will install them first. 

```{r Library Installation and Loading, echo=TRUE, message=FALSE, warning=FALSE}
# Here we store the necessary packages in the necessary_packages variable
necessary_packages <- c("readxl","dplyr","agricolae", "tibble", "ggplot2","ggpubr", "scales","rstatix","ComplexHeatmap","dendextend","vegan","factoextra") 

# Here we check if the necessary packages are installed. If not, R installs them 
if (!requireNamespace(necessary_packages, quietly = TRUE)) {
  install.packages(necessary_packages)
}

# Here, for every package included in the necessary_packages variable, R will load it. The character.only = TRUE ensures that library() will try to interpret the package names as characters (as it needs for it to work) 
for (package in necessary_packages) {
  library(package, character.only = TRUE)
}
```

Step 2: We will load the data. We need two different data formats. 
The first format will be used for the inhibition plots and the descriptive statistics.
The second format will be used for the data visualisation (heatmap, PCA)
  
See here the first format
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(readxl)
head(read_excel("data.xlsx", sheet = 1, col_names = TRUE),12)
```

See here the second format
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(readxl)
head(read_excel("data.xlsx", sheet = 2, col_names = TRUE),12)
```


**Section 1 : Inhibition Plots**

Step 1: Let's import the data for the inhibition plots
  
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Data import
library(readxl)
data_inhibition_plots <- read_excel("data.xlsx", sheet = 1, col_names = TRUE)

# We can change the names of the columns to more R-friendly names
names(data_inhibition_plots) <- c("Exudate", "Strain", "Domain", "AOI", "Mean", "St.Error", "Plant")

#Let's view the data
glimpse(data_inhibition_plots)

# We will store the names of the different root exudates
REs <- as.factor(levels(as.factor(data_inhibition_plots$Exudate)))
REs

```

Step 2: Here we will split the data per root exudate into  different data frames
  
```{r echo=TRUE, message=FALSE, warning=FALSE}
# We will create an empty list that we will fill with all the different data frames 
dataframe_list_REs <- list()

# This for() loop will create the separate data frames 
for (i in 1:summary(levels(REs))[[1]]) {
  data_inh <- data.frame(filter(data_inhibition_plots, Exudate == REs[i]))
  dataframe_name <- paste("df","inhibition",REs[i], sep = "_")
  dataframe_list_REs[[dataframe_name]] <- data_inh
}

# Let's see how many dataframes are inside the list 
nrow(as.matrix(summary(dataframe_list_REs)))

# Let's see which dataframes are inside the list
rownames(as.matrix(summary(dataframe_list_REs)))
```

Step 3: Here we will create the legend order for all the inhibition plots. We need the order to be as follows: Strain A, Strain B, Strain C, Strain D
  
```{r echo=TRUE, message=FALSE, warning=FALSE}
# We will create the legend order 
legend_order <- c("Strain A","Strain B","Strain C","Strain D")
```

Step 4: We will perform Kruskal - Wallis tests with a post - hoc test to differentiate the inhibitory activity of each root exudate against the 4 different AOM strains

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(agricolae)
library(tibble)

# We will create two empty lists to be filled later. One will contain the kruskal - wallis reports and the second one will contain the groupings, in order for them to be assigned over the barplots later 
kruskals <- list()
kruskals_groups <- list()

# The for() loop will perform the test for each data frame in the list we created earlier and will apply the Bonferroni method for p - value adjustment. It will store the reports in the first list. 
# Then it will sort the groupings results by the legend order instead of the value, so the letters can be properly placed above the correct bar in the barplot.
for (i in 1:nrow(as.matrix(summary(dataframe_list_REs)))) {
  kruskal_inhibition <- with(data = dataframe_list_REs[[i]], kruskal(AOI, Strain, p.adj = "bonferroni", group = T, alpha = 0.05))
  kruskal_name <- paste("kruskal",REs[i], sep = "_")
  kruskals[[kruskal_name]] <- kruskal_inhibition
  
  treatment_order <- legend_order
  kruskal_results <- data.frame(kruskal_inhibition$groups)
  kruskal_results <- rownames_to_column(kruskal_results, var = "Strain")
  kruskal_results$Strain <- factor(kruskal_results$Strain, levels = treatment_order)
  kruskal_results <- kruskal_results[order(kruskal_results$Strain), ]
  kruskal_results_name <- paste("kruskal","groups", REs[i], sep = "_")
  kruskals_groups[[kruskal_results_name]] <- kruskal_results
}

# Let's see which dataframes are inside each list
rownames(as.matrix(summary(kruskals)))
rownames(as.matrix(summary(kruskals_groups)))

```

Step 5: Here we will create the inhibition barplots

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(ggplot2)

# We will create an empty list to store all plots
inhibition_barplots <- list()

# The for() loop will create the plots and store them in the list
for (i in 1:nrow(as.matrix(summary(dataframe_list_REs)))) {
  
  # This will sort them correctly in order to avoid mistakes in the plots
  dataframe_list_REs[[i]]$Exudate <- factor(dataframe_list_REs[[i]]$Exudate, levels = legend_order)
      inhibition_plot_temp <- ggplot(filter(dataframe_list_REs[[i]], !is.na(Mean)), aes(x = Strain, y = Mean)) + 
        geom_col(aes(x = Strain, y = Mean, fill = Strain)) + 
        theme_light() + 
        labs(title = paste("Ammonia Oxidation Inhibition % values", REs[i]), x = "AOM Strain", y = "Ammonia Oxidation Inhibition %") + 
        theme(axis.title = element_text(size = 20), axis.text = element_text(size = 20), axis.text.x = element_text(angle = 0), legend.text = element_text(), legend.title = element_text(size = 20), legend.position = "none") + geom_errorbar(aes(ymin = Mean - St.Error, ymax = Mean + St.Error, width = 0.2)) + 
        scale_x_discrete(limits = legend_order) + 
        ylim(-25,125) + 
        geom_text(data = kruskals_groups[[i]], aes(label = groups, y = 110, vjust = -0), size = 14) + 
        theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "inches")) 

  inhibition_plot_name <- paste("inhibition","plot",REs[i],sep = "_")
  inhibition_barplots[inhibition_plot_name] <- list(plot = inhibition_plot_temp)
}

# Let's see one inhibition barplot
inhibition_barplots[37]
```

Step 6: We will generate a .pdf file with all the plots

```{r echo = TRUE, eval=FALSE, warning=FALSE, message=FALSE, results='hide'}

#### Generate the .pdf file ---- 
cairo_pdf("RE_Inhibition plots_all strains.pdf", height = 8, width = 8, onefile = T)
for (i in 1:nrow(as.matrix(summary(inhibition_barplots)))) {
  print(inhibition_barplots[i])
}
dev.off()

graphics.off()

```


**Section 2 : Descriptive statistics**

Step 1: We will use the data in the same format as above. we can clear the environment and re-import the data
  
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Clear environment
rm(list = ls())

# Import the data
library(readxl)
data_statistics <- read_xlsx("data.xlsx", sheet = 1)

# We can remove rows with no AOI data in order not to mess with the analysis
library(dplyr)
data_statistics <- filter(data_statistics, !is.na(AOI))

```

Step 2: We will investigate the distribution of the AOI values in order to decide whether to proceed with parametric or non - parametric tests. If the p-value is lower than the threshold, we debunk the hypothesis that data in each group are normally distributed and perform non-parametric tests
  
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Shapiro - Wilk test for different domains
shapiro_domain_A <- shapiro.test(filter(data_statistics, Domain == "Domain A")$AOI)
shapiro_domain_B <- shapiro.test(filter(data_statistics, Domain == "Domain B")$AOI)

shapiro_domain_A
shapiro_domain_B

  # We can create histograms (indicatevly for the two domains)
AOI_histogram_domain_A <- hist(filter(data_statistics, Domain == "Domain A")$AOI, breaks = 9, main = "Histogram of Ammonia Oxidation Inhibition % values", xlab = "Ammonia Oxidation Inhibition %", col = "skyblue") 

AOI_histogram_domain_B <- hist(filter(data_statistics, Domain == "Domain B")$AOI, breaks = 9, main = "Histogram of Ammonia Oxidation Inhibition % values", xlab = "Ammonia Oxidation Inhibition %", col = "skyblue") 

  # We can also utilise Q-Q plots (indicatevly for the two domains)
qq_plot_domain_A <- qqnorm(filter(data_statistics, Domain == "Domain A")$AOI, main = "Q-Q plot of Ammonia Oxidation Inhibition % values", col = "skyblue4")

qq_plot_domain_B <- qqnorm(filter(data_statistics, Domain == "Domain B")$AOI, main = "Q-Q plot of Ammonia Oxidation Inhibition % values", col = "skyblue4")

# Shapiro - Wilk test for different strains
shapiro_strain_A <- shapiro.test(filter(data_statistics, Strain == "Strain A")$AOI)
shapiro_strain_B <- shapiro.test(filter(data_statistics, Strain == "Strain B")$AOI)
shapiro_strain_C <- shapiro.test(filter(data_statistics, Strain == "Strain C")$AOI)
shapiro_strain_D <- shapiro.test(filter(data_statistics, Strain == "Strain D")$AOI)

shapiro_strain_A 
shapiro_strain_B 
shapiro_strain_C 
shapiro_strain_D 

# Shapiro - Wilk test for different plants
shapiro_plant_A <- shapiro.test(filter(data_statistics, Plant == "Plant A")$AOI)
shapiro_plant_B <- shapiro.test(filter(data_statistics, Plant == "Plant B")$AOI)

shapiro_plant_A 
shapiro_plant_B 

```

Step 3: We will first compare the mean AOI values between the 2 Domains of AOM strains 
  
```{r echo=TRUE, message=FALSE, warning=FALSE}
# We will group the data based on the domain and investigate the means
library(dplyr)
data_grouped_domain <- group_by(filter(data_statistics, !is.na(AOI)), Domain)
summarise(data_grouped_domain, mean_AOI = mean(AOI))

# We will compute a Wilcoxon test 
library(rstatix)
wilcox_test(data_statistics, AOI ~ Domain)

# We will generate a box - plot and incorporate the result of the Wilcoxon test
library(ggplot2)
library(ggpubr)

AOIs_boxplot_domain <- ggplot(data_statistics, aes(x = Domain, y = AOI, col = Domain)) +
  geom_boxplot(aes(colour = Domain)) + 
  theme_classic() + 
  labs(title = "Ammonia Oxidation Inhibition %", subtitle = "Across the different AOM Domains", x = "AOM Domain", y = "Ammonia Oxidation Inhibition % (AOI%)") + 
  theme(axis.title = element_text(size = 20), axis.text.y = element_text(size = 20), axis.text.x = element_text(size = 14, angle = 0, vjust = 0), legend.text = element_text(size = 20), legend.title = element_text(size = 20), legend.position = "none") + 
  ylim(-25,140) + 
  geom_text(x = -Inf, y = -Inf, label = paste("Wilcoxon test: W - statistic =", round(wilcox_test(data_statistics, AOI ~ Domain)$statistic,2), ", p value =", wilcox_test(data_statistics, AOI ~ Domain)$p), vjust = -1, hjust = -0.05, colour = "black", size = 4) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("Domain A","Domain B")))

AOIs_boxplot_domain
```

```{r echo = TRUE, eval=FALSE, warning=FALSE, message=FALSE, results='hide'}
cairo_pdf("AOI values boxplot_domain.pdf", height = 8, width = 7)
print(AOIs_boxplot_domain)
dev.off()

```

Step 4: We will now compare the mean AOI values between the 4 AOM strains 
  
```{r echo=TRUE, message=FALSE, warning=FALSE}
# We will group the data based on the domain and investigate the means
library(dplyr)
data_grouped_strain <- group_by(filter(data_statistics, !is.na(AOI)), Strain)
summarise(data_grouped_strain, mean_AOI = mean(AOI))

# We will compute a Kruskal - Wallis test and sort the groups as we did above for the inhibition barplots, so they appear correctly over the corresponding box 

library(rstatix)
kruskal_test(data_statistics, AOI ~ Strain)

library(agricolae)
kruskal_AOIs_Strain <- with(data = data_statistics, kruskal(AOI, Strain, p.adj = "bonferroni", group = T, alpha = 0.05))
kruskal_AOIs_Strain$groups

library(tibble)
legend_order <- c("Strain A","Strain B","Strain C","Strain D")

kruskal_results <- data.frame(kruskal_AOIs_Strain$groups)
kruskal_results <- rownames_to_column(kruskal_results, var = "Strain")
strain_order <- legend_order
kruskal_results$Strain <- factor(kruskal_results$Strain, levels = strain_order)
kruskal_results <- kruskal_results[order(kruskal_results$Strain), ]
view(kruskal_results)

# We will generate a box - plot and incorporate the result of the Kruskal - Wallis test
library(ggplot2)
library(ggpubr)

AOIs_boxplot_strains <- ggplot(data_statistics, aes(x = Strain, y = AOI, col = Strain)) +
  geom_boxplot(aes(colour = Strain)) + 
  theme_classic() + 
  labs(title = "Ammonia Oxidation Inhibition %", subtitle = "Across the different AOM Strains", x = "AOM Strain", y = "Ammonia Oxidation Inhibition % (AOI%)") + 
  theme(axis.title = element_text(size = 20), axis.text.y = element_text(size = 20), axis.text.x = element_text(size = 14, angle = 0, vjust = 0), legend.text = element_text(size = 20), legend.title = element_text(size = 20), legend.position = "none") + 
  ylim(-25,140) + 
  geom_text(x = -Inf, y = -Inf, label = paste("Kruskal - Wallis: Ï‡\u00B2 =", round(kruskal_test(data_statistics, AOI ~ Strain)$statistic,2), ", p value =", kruskal_test(data_statistics, AOI ~ Strain)$p), vjust = -1, hjust = -0.05, colour = "black", size = 4) +
  geom_text(data = data.frame(kruskal_results), aes(label = groups, y = data_Normalised_means$Mean), y = 115, size = 8, colour = "black")

AOIs_boxplot_strains
```

```{r echo = TRUE, eval=FALSE, warning=FALSE, message=FALSE, results='hide'}
cairo_pdf("AOI values boxplot_strains.pdf", height = 8, width = 7)
print(AOIs_boxplot_strains)
dev.off()

```

Step 5: Finally, we will compare the mean AOI values produced by the root exudates of the genotyes from the 2 different plant species 
  
```{r echo=TRUE, message=FALSE, warning=FALSE}
# We will group the data based on the domain and investigate the means
library(dplyr)
data_grouped_plant <- group_by(filter(data_statistics, !is.na(AOI)), Plant)
summarise(data_grouped_plant, mean_AOI = mean(AOI))

# We will compute a Wilcoxon test 
library(rstatix)
wilcox_test(data_statistics, AOI ~ Plant)

# We will generate a box - plot and incorporate the result of the Wilcoxon test
library(ggplot2)
library(ggpubr)

AOIs_boxplot_plant <- ggplot(data_statistics, aes(x = Plant, y = AOI, col = Plant)) +
  geom_boxplot(aes(colour = Plant)) + 
  theme_classic() + 
  labs(title = "Ammonia Oxidation Inhibition %", subtitle = "Produced by exudates from the different plants", x = "Plant species", y = "Ammonia Oxidation Inhibition % (AOI%)") + 
  theme(axis.title = element_text(size = 20), axis.text.y = element_text(size = 20), axis.text.x = element_text(size = 14, angle = 0, vjust = 0), legend.text = element_text(size = 20), legend.title = element_text(size = 20), legend.position = "none") + 
  ylim(-25,140) + 
  geom_text(x = -Inf, y = -Inf, label = paste("Wilcoxon test: W - statistic =", round(wilcox_test(data_statistics, AOI ~ Plant)$statistic,2), ", p value =", wilcox_test(data_statistics, AOI ~ Plant)$p), vjust = -1, hjust = -0.05, colour = "black", size = 4) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = list(c("Plant A","Plant B")))

AOIs_boxplot_plant
```

```{r echo = TRUE, eval=FALSE, warning=FALSE, message=FALSE, results='hide'}
cairo_pdf("AOI values boxplot_plant.pdf", height = 8, width = 7)
print(AOIs_boxplot_plant)
dev.off()

```


**Section 3 : Data Visualisation - Heatmap**

Step 1: We will use the data in the second format. We can clear the environment and import the data
  
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Clear environment
rm(list = ls())

# Import the data
library(readxl)
data_heatmap <- read_xlsx("data.xlsx", sheet = 2)

# Let's view the data
data_heatmap

```

Step 2: The data need some pre-processing before they are valid for the package we will use to generate the heatmap 
  
```{r echo=TRUE, message=FALSE, warning=FALSE}
# We will remove the exudate names and store them in a separate vector
library(dplyr)
root_exudates <- data_heatmap$Exudate
data_AOI <- select(data_heatmap, -Exudate)

# We will do the same with the Plant variable
data_AOI <- select(data_AOI, -Plant)

# We will turn the data frame into a matrix, and we will put the root exudate names as row names
data_AOI_mat <- as.matrix(data_AOI)
col_to_row_names <- root_exudates
rownames(data_AOI_mat) <- col_to_row_names

# Let's view the data now
data_AOI_mat

```

Step 3: We will create an initial heatmap 
  
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(ComplexHeatmap)
library(dendextend)

# These commands set the dendrogram for the rows (Root Exudates) and the columns (AOM Strains) of the heatmap, including the type of clustering (here hierarchical) and the type of distances calculated (here Euclidean distances). We also set the different colours for the different clusters. Here, for the initial heatmap we will set it to one colour
row_dend <- as.dendrogram(hclust(dist(data_AOI_mat, method = "euclidean"), method = "complete"))
row_dend <- colour_branches(row_dend, k = 1) 
column_dend <- as.dendrogram(hclust(dist(data_AOI_mat, method = "euclidean"), method = "complete"))
column_dend <- colour_branches(column_dend, k = 1)

RE_AOI_heatmap <- Heatmap(data_AOI_mat, name = "Ammonia Oxidation Inhibition %", 
        column_title = "AOM strains", column_title_side = "bottom", column_title_gp = gpar(fontsize = 16), column_dend_side = "bottom", column_names_gp = gpar(fontsize = 10),
        row_title = "Root Exudate", row_names_side = "left",  row_names_gp = gpar(fontsize = 12),   row_dend_width = unit(10,"mm"), row_title_gp = gpar(fontsize = 16),
        column_title_rot = 0, column_names_rot = 0, column_names_centered = T, 
        col = c("lightblue3", "lightblue1", "yellow", "orange", "orange3", "red"), 
        cluster_rows = row_dend, row_dend_gp = gpar(font = "bold"),
        border = F, 
        column_km = 1, 
        show_column_dend = T,
        width = ncol(data_AOI_mat)*unit(25,"mm"), height = nrow(data_AOI_mat)*unit(4.5,"mm"), 
        cell_fun = function(j, i, x, y, width, height, fill) {
  grid.text(sprintf("%.2f", data_AOI_mat[i, j]), x, y, gp = gpar(fontsize = 10))
})

RE_AOI_heatmap
```

```{r echo = TRUE, eval=FALSE, warning=FALSE, message=FALSE, results='hide'}
cairo_pdf("Heatmap_initial.pdf", width = 8, height = 16)
RE_AOI_heatmap
dev.off()

```

Step 4: Based on this initial heatmap we will designate the clustering groups and also prepare an annotation on the right side indicating the plant species from which the root exudates originate
  
```{r echo=TRUE, message=FALSE, warning=FALSE}
# We will extract the clusters
row_order_clustered <- row_order(RE_AOI_heatmap)

# We will bind the matrix we use for the heatmap with the clustered order of the root exudates
data_clustered <- cbind(data_AOI_mat, row_order_clustered)

# We will create a data frame and go through some transformations. By the end, we will have sorted the root exudates with the order they appear on the heatmap (top to bottom). And they are accompanied by the plant they originate from 
data_plant <- data.frame(cbind(data_heatmap$Exudate, data_heatmap$Plant))
plant_origin_ordered <- cbind(data_plant, seq(1,77,1))
colnames(plant_origin_ordered) <- c("Exudate","Plant","Number")
plant_origin_ordered <- mutate(plant_origin_ordered, Number = as.numeric(Number))
plant_origin_ordered$Number <- factor(plant_origin_ordered$Number, levels = row_order_clustered)
plant_origin_ordered <- plant_origin_ordered[order(plant_origin_ordered$Number), ]
view(plant_origin_ordered)

# We will extract the plant order to prepare the annotation on the right side of the heatmap
plant_order <- plant_origin_ordered$Plant
RE_order <- plant_origin_ordered$Exudate
```

Step 5: Now we will sort the matrix with the AOI values based on the order that the exudates appear on the heatmap 

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tibble)
data_AOI_origin <- data_AOI_mat
data_AOI_origin <- data.frame(data_AOI_origin)
data_AOI_origin <- rownames_to_column(data_AOI_origin, var = "Exudate")

data_AOI_origin$Exudate <- factor(data_AOI_origin$Exudate, levels = RE_order)
data_AOI_origin <- data_AOI_origin[order(data_AOI_origin$Exudate), ]
data_AOI_origin <- cbind(seq(1,77,1), data_AOI_origin)
rownames(data_AOI_origin) <- data_AOI_origin$"seq(1, 77, 1)"
rownames(data_AOI_origin) <- data_AOI_origin$Exudate
data_AOI_origin <- data_AOI_origin[,-c(1:2)]

data_AOI_origin_mat <- as.matrix(data_AOI_origin)

data_AOI_origin_mat
```

Step 6: We will designate the annotation and prepare the final heatmap 

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(ComplexHeatmap)
library(dendextend)

# we will designate the type of annotation and the colours of the different plant species 
row_ha_Origin = rowAnnotation("Plant species" = plant_order, col = list("Plant species" = c("Plant A" = "orange", "Plant B" = "green2")))

# We will re - designate the cluster number based on the initial heatmap
row_dend <- as.dendrogram(hclust(dist(data_AOI_mat, method = "euclidean"), method = "complete"))
row_dend <- colour_branches(row_dend, k = 13) 
column_dend <- as.dendrogram(hclust(dist(data_AOI_mat, method = "euclidean"), method = "complete"))
column_dend <- colour_branches(column_dend, k = 1)

RE_AOI_heatmap_final <- Heatmap(data_AOI_origin_mat, name = "Ammonia Oxidation Inhibition %",  right_annotation = row_ha_Origin,
        column_title = "AOM strains", column_title_side = "bottom", column_title_gp = gpar(fontsize = 16), column_dend_side = "bottom", column_names_gp = gpar(fontsize = 10),
        row_title = "Root Exudate", row_names_side = "left",  row_names_gp = gpar(fontsize = 12),   row_dend_width = unit(10,"mm"), row_title_gp = gpar(fontsize = 16),
        column_title_rot = 0, column_names_rot = 0, column_names_centered = T, 
        col = c("lightblue3", "lightblue1", "yellow", "orange", "orange3", "red"), 
        cluster_rows = row_dend, row_dend_gp = gpar(font = "bold"),
        border = F, 
        column_km = 1,
        show_column_dend = T,
        width = ncol(data_AOI_mat)*unit(25,"mm"), height = nrow(data_AOI_mat)*unit(4.5,"mm"), 
        cell_fun = function(j, i, x, y, width, height, fill) {
  grid.text(sprintf("%.2f", data_AOI_mat[i, j]), x, y, gp = gpar(fontsize = 10))
})

RE_AOI_heatmap_final
```

```{r echo = TRUE, eval=FALSE, warning=FALSE, message=FALSE, results='hide'}
cairo_pdf("Heatmap_final_annotated.pdf", width = 8, height = 16)
RE_AOI_heatmap_final
dev.off()

```


**Section 4 : Data Visualisation - PCA**

Step 1: We will use the data in the second format. We can clear the environment and 
re - import the data
  
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Clear environment
rm(list = ls())

# Import the data
library(readxl)
data_PCA <- read_xlsx("data.xlsx", sheet = 2)

# Let's view the data
data_PCA

```

Step 2: The data need some pre - processing
```{r echo=TRUE, message=FALSE, warning=FALSE}
# We need to keep only the AOI values
library(dplyr)
data_mat <- select(data_PCA, -c(Exudate, Plant))

# We need to scale our data
library(vegan)
data_standardised <- scale(data_mat, center = TRUE, scale = TRUE) 

# We will use the root exudates as row names in the matrix
rownames(data_standardised) <- data_PCA$Exudate

# Let's inspect the matrix
data_standardised

```

Step 3: Let's perform the PCA 

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(factoextra)

# We compute the PCA and extract the data 
pca <- prcomp(data_standardised, scale. = TRUE)
pca_data <- as.data.frame(pca$x)

# We can inspect the Eigenvalues to see if our principal components explain a sufficient amount of the original variation
fviz_eig(pca)
get_eigenvalue(pca)

# We will enter also the Plant variable in order to perform some clustering later
pca_data <- mutate(pca_data, Plant = data_PCA$Plant)

# We will plot first a simple PCA
REs_PCA_plot <- fviz_pca_ind(pca,
                             habillage = as.factor(pca_data$Plant), 
                             addEllipses = F,
                             repel = T, geom = "text",
                             legend.title = "Plant",
                             palette = c("green2", "orange"))
REs_PCA_plot
```

```{r echo = TRUE, eval=FALSE, warning=FALSE, message=FALSE, results='hide'}
cairo_pdf("PCA_simple.pdf", width = 8, height = 8)
REs_PCA_plot
dev.off()
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
# We will now investigate the contribution and correlation between the different variables (here the 4 AOM Strains)
REs_strain_correlation_plot <- fviz_pca_var(pca, 
                                            col.var = "contrib", 
                                            gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
                                            repel = TRUE, 
                                            title = "AOM strains correlation")
REs_strain_correlation_plot
```

```{r echo = TRUE, eval=FALSE, warning=FALSE, message=FALSE, results='hide'}
cairo_pdf("PCA_variables.pdf", width = 8, height = 8)
REs_strain_correlation_plot
dev.off()
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
# We can try and investigate any clustering due to the plant species, so we will generate the PCA plot, insert ellipses and complement with a PERMANOVA
REs_PCA_plot_elipses <- fviz_pca_ind(pca,
                                     habillage = as.factor(pca_data$Plant), 
                                     addEllipses = T,
                                     repel = T, geom = "text",
                                     legend.title = "Plant",
                                     palette = c("green2", "orange"))
REs_PCA_plot_elipses

data_standardised_permanova <- cbind(data.frame(data_standardised), data_PCA$Plant)
adonis2(data_standardised_permanova[,1:4] ~ data_PCA$Plant, method= 'euc')
```

```{r echo = TRUE, eval=FALSE, warning=FALSE, message=FALSE, results='hide'}
cairo_pdf("PCA_ellipses.pdf", width = 8, height = 8)
REs_PCA_plot_elipses
dev.off()
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Finally we can generate a bi-plot with the ellipses and the variable vectors
REs_biplot <- fviz_pca_biplot(pca,
                              habillage = as.factor(pca_data$Plant),
                              addEllipses = T, ellipse.level = 0.95,
                              legend.title = "Plant",
                              repel = T, 
                              geom = "text",
                              palette = c("green2", "orange"))
REs_biplot
```

```{r echo = TRUE, eval=FALSE, warning=FALSE, message=FALSE, results='hide'}
cairo_pdf("PCA_biplot.pdf", width = 8, height = 8)
print(REs_biplot)
dev.off()

```

This is the end. We have analysed the data, starting investigating the mean activity of all exudates against the 2 domains and the 4 strains of AOM, as well as the effect of the 2 different plant species. Then, through the inhibition plots, we investigated the variation in the activity of each exudate separately. Finally, we visualised our data, using a heatmap and a PCA, to discover distinct inhibitory activity patterns.  

Thanks a lot for using this code! 

Remember to always cite the author of the code you use!

And buy a programmer a beer ðŸº ! They really need one!
